---
title: "Project 3"
execute: 
  warning: false
  message: true
---

Research Question: 
Using the "College tuition, diversity, and pay" data set, I will be analyzing the differences in cost/tuition between public and private schools. I will filter out the data in order to make a fair comparison between the two categories (e.g. schools in the same region), then compare the difference between their costs. 
Here is a link to my data set: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-10/readme.md

Here are my research questions:
Alternative claim: Tuition rates are higher for private universities than for public universities 
Null claim: Tuition rates do not differ between public and private universities.

```{r}
#| echo: true
library(tidyverse)
library(purrr)

#tuesdata <- tidytuesdayR::tt_load('2020-03-10')
#tuesdata <- tidytuesdayR::tt_load(2020, week = 11)


#tuition_cost <- tuesdata$tuition_cost
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-03-10/tuition_cost.csv')
```
I am only interested in the total cost for students attending schools in-state, in California. This way, the cost-of-living is fairly looked at (since comparing Californiaa schools with Maine schools, for example, would not be fair). 
```{r}
library(dplyr)

# Sorting my data:
CA_schools <- tuition_cost |>
  select(name, state_code, type, in_state_tuition) |>
  filter(state_code == "CA") |>
  filter(type %in% c("Public", "Private"))

CA_schools 
# public schools
#public_schools <- tuition_cost |>
#  select(name, state_code, type, in_state_tuition) |>
#  filter(type == "Public") |>
#  filter(state_code == "CA")

#public_mean <- public_schools |>
#  summarize(mean_cost = mean(in_state_tuition, na.rm = TRUE))

#145 total public schools
#public_schools
#round(public_mean)


# private schools
#private_schools <- tuition_cost |>
#  select(name, state_code, type, in_state_tuition) |>
#  filter(type == "Private") |>
#  filter(state_code == "CA") 

#private_mean <- private_schools |>
#  summarize(mean_cost = mean(in_state_tuition, na.rm = TRUE))

# 94 total private schools
#private_schools
#round(private_mean)
```

```{r}
# My function to simulate a random sampling of school's in-state tuition

# average
avg_cost <- CA_schools |>
  group_by(type) |>
  summarize(avg_cost = mean(in_state_tuition))

avg_diff <- CA_schools |> 
  group_by(type) |> 
  summarize(avg_cost = mean(in_state_tuition)) |> 
  summarize(ave_diff = diff(avg_cost))

avg_diff_val <- abs(avg_diff$ave_diff[1])

avg_diff_val
#########################################################

perm_data <- function(rep, data){
  CA_schools |>
    select(type, in_state_tuition) |> 
    mutate(perm_type = sample(type)) |> 
    group_by(perm_type) |> 
    dplyr::summarize(perm_cost = mean(in_state_tuition, na.rm = TRUE)) |>
    summarize(diff_perm = diff(perm_cost), rep = rep)
}

#map(1:10, perm_data, data = avg_diff) |> 
#  list_rbind()

set.seed(1993)
# 1000 iterations
num_exper <- 1000
simulated_differences <- map(1:num_exper, perm_data, data = avg_diff) |> 
  list_rbind()

simulated_differences
#simulated_differences |>
#  ggplot(aes(x = diff_perm)) +   # Use diff_perm for the x-axis
#  geom_histogram(binwidth = 500, fill = "pink", color = "white") +   # Customize histogram
#  labs(
#    x = "Difference of Mean Cost of Attendance",
 #   title = "Sampling Distribution of Permuted Differences"
  #)


simulated_differences |> 
  data.frame() |> 
  ggplot(aes(x = diff_perm)) + 
  geom_histogram(binwidth = 500, fill = "pink", color = "white") + 
  geom_vline(xintercept = avg_diff_val, color = "red") + 
  labs(x = "The difference of mean cost of attendance betwen Public and Private Universities",
       title = "Sampling Distribution of Permuted Differences")
  
perm_stats |> 
  summarize(p_val_ave = mean(diff_perm > avg_diff_val))
perm_stats

#########################################################


# permute on private / publuc
# generates a single random sample
null_sample <- CA_schools |>
  mutate(perm_type = sample(type)) |>
  group_by(perm_type) |>
  dplyr::summarize(perm_cost = mean(in_state_tuition, na.rm = TRUE)) |>
  summarize(diff_perm = diff(perm_cost))

# look at difference between average, and this code ^^^^

# generate a null distribution
perm_data <- function(rep, sample_size, avg_diff) {
  # Sample the data
  sampled_data <- CA_schools |> sample_n(sample_size)
  
  # Permute types and calculate permuted difference
  perm_results <- sampled_data |>
    mutate(perm_type = sample(type)) |>
    group_by(perm_type) |>
    dplyr::summarize(perm_cost = mean(in_state_tuition, na.rm = TRUE)) |>
    summarize(diff_perm = diff(perm_cost))
  
  # Calculate the difference between permuted and average difference
  perm_results <- perm_results |>
    mutate(diff_from_avg = diff_perm - avg_diff, rep = rep)
  
  return(perm_results)
}
# make a histogram of these differences
# we don't want both, we want difference
set.seed(1993)

# 1000 iterations
num_exper <- 1000
#simulated_differences <- map_dbl(1:num_exper, random_sample, sample_size = 100)
#simulated_differences <- map_dbl(1:num_exper, perm_data, sample_size = 100)
simulated_results <- map_df(1:num_exper, ~perm_data(rep = .x, sample_size = 100, avg_diff = avg_diff))

simulated_differences |> 
  data.frame() |> 
  ggplot(aes(x = simulated_differences)) + 
  geom_histogram(binwidth = 500, fill = "pink", color = "white") + 
  geom_vline(aes(xintercept = observed_diff), color = "red") +
  labs(x = "The difference of mean cost of attendance betwen Public and Private Universities",
       title = "Sampling distribution when null hypothesis is true")
  

p_value <- mean(abs(simulated_differences) >= abs(observed_diff))
p_value

#random_sample <- function(rep, sample_size) {
  # grabs a slice of schools at 
#  sampled_schools <- CA_schools |>
#    sample_n(size = sample_size, replace = FALSE)
  
  # mean for public schools
#  mean_public_tuition <- sampled_schools |>
#    filter(type == "Public") |>
#    summarize(mean_tuition = mean(in_state_tuition, na.rm = TRUE)) |>
#    pull(mean_tuition)
  
  # mean for private schools
#  mean_private_tuition <- sampled_schools |>
#    filter(type == "Private") |>
  #  summarize(mean_tuition = mean(in_state_tuition, na.rm = TRUE)) |>
  #  pull(mean_tuition)
  
  #difference <- mean_private_tuition - mean_public_tuition
  #return(difference)
#}

# Calculate the difference in the previous observed means
#observed_diff <- as.integer(private_mean - public_mean) 
#observed_diff <- as.integer(avg_sample - null_sample) 



```
This plot shows us the results from randomly shuffling all of the tuition costs between public and private universities, and then measuring their mean difference.

From these data, the observed differences seem to be consistent with the distribution of differences in the null sampling distribution.

There is no strong evidence to reject the null hypothesis.

We can claim that the tuition rates are not higher for private universities than for public universities (p-value = .502).

We achieved this result by filtering for California schools, and then calculating the average mean between private and public schools in California. We then created a random sampling by cutting a slice of elements from the data set of all Californian schools, and randomly shuffling their tuition values. We ran this over 2000 iterations, and found that there is strong evidence for the null hypothesis in this case.


